---
title: "CASE 4 - Teaching"
author: Michael Bigler, Tania Loureiro, William Elkiess, Dakota Cuellar
  and Ilyana El Mendili, Liam Phan
date: "2023-03-10"
output: 
  rmdformats::downcute:
    code_folding: hide
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE,comment = FALSE, error = FALSE,options(scipen=999))
rm(list = ls())

```

# Packages

```{r packages}
library(knitr) # html tables
library(gtsummary)
```

# Data Loading

```{r data loading}
teaching <- readxl::read_excel('Bteaching.xls')
teaching <- data.frame(teaching)
teaching$GROUP <- factor(teaching$GROUP, level = c('C', 'E'), labels = c('Control', 'Experiment'))
teaching$GENDER <- factor(teaching$GENDER, level = c('M', 'F'), labels =  c('Male', 'Female'))
teaching$PREF <- factor(teaching$PREF, level = c('A', 'N', 'T', 'K', 'V'), labels = c('A', 'N', 'T', 'K', 'V')) # get understandable lables
teaching$SLEVEL <- factor(teaching$SLEVEL, level = c('E', 'H'), labels = c('Elementary', 'Highschool'))
```

# Data exploratory

```{r exploratory}
kable(summary(teaching))
```

```{r}
teaching$index = 1:nrow(teaching)
tsibble::tsibble(teaching, index = index)

table1 <- tbl_summary(
  teaching,
  by = GROUP, # split table by group
  missing = "no" # don't list missing data separately
) |> 
  add_n() |>  # add column with total number of non-missing observations
  add_p() |>  # test for a difference between groups
  modify_header(label = "**Variable**") |> # update the column header
  bold_labels() 
table1
```

```{r}
tbl_strata <-
  teaching |> 
  tbl_strata(
    strata = GROUP,
    .tbl_fun =
      ~ .x |>
        tbl_summary(by = SESSION, missing = "no") |>
      add_n() |>  # add column with total number of non-missing observations
      add_p() |>  # test for a difference between groups
      modify_header(label = "**Variable**") |> # update the column header
      bold_labels() ,
    .header = "**{strata}**, N = {n}"
  )
tbl_strata
```

## Dummy encoding

```{r}
library(fastDummies)
teaching <- dummy_columns(teaching, select_columns = c('GROUP', 'SESSION', 'PREF', 'GENDER', 'SLEVEL'), remove_selected_columns = TRUE, remove_first_dummy = TRUE)
```

```{r}
library(dplyr)
teaching$DIFF <- teaching$POST - teaching$PRE
mod <- lm(DIFF ~ ., data = subset(teaching, select = -c(PRE, POST)))
summary(mod)
plot(mod)


source(file="RFn_Plot-lmSim.R")
plot.lmSim(mod)
```

I guess we can assume normality as qqplot in bootstrap.

```{r}
library(regclass)
VIF(mod)
```

Shows which variables have higher importance

# Selection of variables 

[selections](https://cran.r-project.org/web/packages/olsrr/vignettes/variable_selection.html)
[more selections and lasso regression](https://rpubs.com/fractalbass/601525)


## regsubsets

```{r}
library(leaps)
models <- regsubsets(DIFF~., data = subset(teaching, select = -c(PRE, POST)), nvmax = 10,
                     method = "seqrep")
summary(models)

# http://www.sthda.com/english/articles/37-model-selection-essentials-in-r/155-best-subsets-regression-essentials-in-r/
```
## forward 

```{r}
mod4 <- lm(DIFF ~ 1, subset(teaching, select = -c(PRE, POST)))
step(mod4, direction='forward', scope=formula(lm(DIFF ~ ., data = subset(teaching, select = -c(PRE, POST)))))
```

## backward 

```{r}
s <- step(mod, direction = 'backward')
s
mod2 <- lm(DIFF ~ ATT + GROUP_Experiment + SESSION_S2 + PREF_N + PREF_K + PREF_V, subset(teaching, select = -c(PRE, POST)))

summary(mod2)
```

Selects the best model backwards which makes some variables significant. 

Now a model which would make sense

```{r}
# model due to sense
mod3 <- lm(DIFF ~ ATT + GROUP_Experiment + PREF_N + PREF_K + 
    PREF_V + PREF_T, subset(teaching, select = -c(PRE, POST)))

summary(mod3)
```


```{r include=FALSE}
library(readxl)
BB <- read_excel("Bteaching.xls")
bt <- read_excel("Bteaching.xls")
data <- read_excel("Bteaching.xls")
```

$$
We\ build\ a\ GLM\ model:\\ \\POST\ = GROUP + PREF + PRE
$$

```{r}
# Convert GROUP and PREF to factors
#data$GROUP <- factor(data$GROUP)
#data$PREF <- factor(data$PREF)
data <- read_excel("Bteaching.xls")
library(ggplot2)
data$GENDER <- ifelse(data$GENDER == "M", 0, 1)
data$GROUP <- ifelse(data$GROUP == "C", 0, 1)
data$SLEVEL <- ifelse(data$SLEVEL == "E", 0, 1)
data$SESSION <- ifelse(data$SESSION == "S1", 1, 
                        ifelse(data$SESSION == "S2", 2, 
                               ifelse(data$SESSION == "S3", 3, 
                                      ifelse(data$SESSION == "S4", 4, NA))))
data$PREF <- ifelse(data$PREF == "A", 1, 
                        ifelse(data$PREF == "K", 2, 
                               ifelse(data$PREF == "N", 3, 
                                      ifelse(data$PREF == "T", 4,
                                             ifelse(data$PREF == "V", 5, NA)))))

colnames(data)[1] <- "Group"
colnames(data)[2] <- "Session"
colnames(data)[3] <- "Preference"
colnames(data)[4] <- "Gender"
colnames(data)[5] <- "School Level"
colnames(data)[6] <- "Pre Test Score"
colnames(data)[7] <- "Post Test Score"
colnames(data)[8] <- "Attention Test Score"
require(corrplot)

#2B3433, #2F4F4E, #0D8387, #CBE4DE
#("#FF0000", "#FFFFFF", "#0000FF")
cor_mat <- cor(data[, 7], data[, -7], method = "spearman")
col <- colorRampPalette(c("#2B3433", "#FFFFFF", "#0D8387"))(25)
corrplot(cor_mat, col = col, tl.col = "black",
         tl.srt = 40, tl.cex = 1.2,
         title = "Correlation Matrix", mar = c(0, 0, 1, 0))
#I have this corrplot i want to order the correlation circles from highest to lowest correlation
# Select the 7th column of the data and compute the correlation matrix


```

```{r}
data <- read_excel("Bteaching.xls")

data$SLEVEL <- factor(data$SLEVEL, level = c('E', 'H'), labels = c('Elementary', 'Highschool'))
data$PREF <- factor(data$PREF, level = c('A', 'N', 'T', 'K', 'V'), labels = c('A', 'N', 'T', 'K', 'V'))
data$GROUP <- factor(data$GROUP)


modelGLMAAA <- glm(formula = POST ~ GROUP + PRE, data = data, family =gaussian(link = "identity"))
summary(modelGLMAAA)

```


```{r}
data <- read_excel("Bteaching.xls")
data$GENDER <- ifelse(data$GENDER == "M", 0, 1)
data$GROUP <- ifelse(data$GROUP == "C", 0, 1)
data$SESSION <- ifelse(data$SESSION == "S1", 1, 
                        ifelse(data$SESSION == "S2", 2, 
                               ifelse(data$SESSION == "S3", 3, 
                                      ifelse(data$SESSION == "S4", 4, NA))))
bn <- subset(data, PRE < 56)
bf <- subset(data, PRE > 57)
bn$SLEVEL <- factor(bn$SLEVEL, level = c('E', 'H'), labels = c('Elementary', 'Highschool'))
bn$PREF <- factor(bn$PREF, level = c('A', 'N', 'T', 'K', 'V'), labels = c('A', 'N', 'T', 'K', 'V'))
bn$GROUP <- factor(bn$GROUP)


bf$SLEVEL <- factor(bf$SLEVEL, level = c('E', 'H'), labels = c('Elementary', 'Highschool'))
bf$PREF <- factor(bf$PREF, level = c('A', 'N', 'T', 'K', 'V'), labels = c('A', 'N', 'T', 'K', 'V'))
bf$GROUP <- factor(bf$GROUP)

modelGLMbn <- glm(formula = POST ~  GROUP + PRE , data = bn, family =gaussian(link = "identity"))
summary(modelGLMbn)
modelGLMbf <- glm(formula = POST ~  GROUP + PRE , data = bf, family =gaussian(link = "identity"))
summary(modelGLMbf)
```

```{r}
data <- read_excel("Bteaching.xls")
data$GENDER <- ifelse(data$GENDER == "M", 0, 1)
data$GROUP <- ifelse(data$GROUP == "C", 0, 1)
data$SESSION <- ifelse(data$SESSION == "S1", 1, 
                        ifelse(data$SESSION == "S2", 2, 
                               ifelse(data$SESSION == "S3", 3, 
                                      ifelse(data$SESSION == "S4", 4, NA))))
bE <- subset(data, SLEVEL == "E")
bH <- subset(data, SLEVEL == "H")
bE$SLEVEL <- factor(bE$SLEVEL, level = c('E', 'H'), labels = c('Elementary', 'Highschool'))
bE$PREF <- factor(bE$PREF, level = c('A', 'N', 'T', 'K', 'V'), labels = c('A', 'N', 'T', 'K', 'V'))
bE$GROUP <- factor(bE$GROUP)


bH$SLEVEL <- factor(bH$SLEVEL, level = c('E', 'H'), labels = c('Elementary', 'Highschool'))
bH$PREF <- factor(bH$PREF, level = c('A', 'N', 'T', 'K', 'V'), labels = c('A', 'N', 'T', 'K', 'V'))
bH$GROUP <- factor(bH$GROUP)

modelGLMbE <- glm(formula = POST ~  GROUP + PRE, data = bE, family =gaussian(link = "identity"))
summary(modelGLMbE)
modelGLMbH<- glm(formula = POST ~  GROUP + PRE, data = bH, family =gaussian(link = "identity"))
summary(modelGLMbH)
```




```{r}
data <- read_excel("Bteaching.xls")

data$GENDER <- factor(data$GENDER)
data$GROUP <- factor(data$GROUP)
data$SLEVEL <- factor(data$SLEVEL)
data$SESSION <- factor(data$SESSION)
data$PREF <- factor(data$PREF)
#data$PREF1A <- ifelse(data$PREF == "A", 1, 10)
#data$PREF2K <- ifelse(data$PREF == "K", 2, 10)
#data$PREF3N <- ifelse(data$PREF == "N", 3, 10)
#data$PREF5V <- ifelse(data$PREF == "V", 5, 10)
#data$PREF4T <- ifelse(data$PREF == "T", 4, 10)

modelGLM2 <- glm(POST ~  PRE + PREF + GENDER, data = data, family =gaussian(link = "identity"))
summary(modelGLM2)
plot(modelGLM2)
```


```{r}
library(ggplot2)
btc <- subset(BB, GROUP == "C")
bte <- subset(BB, GROUP == "E")

# Combine the data into a data frame
dataCE <- data.frame(group = c(rep("Control", length(btc$POST)), rep("Experiment", length(bte$POST))), value = c(btc$POST, bte$POST))

# Create the box plot graph
ggplot(dataCE, aes(x = group, y = value, fill = group)) + geom_boxplot() + theme_classic()

# Combine the data into a data frame
dataCE <- data.frame(group = c(rep("Control", length(btc$POST)), rep("Experiment", length(bte$POST))), value = c(btc$POST, bte$POST))


# Data frame for each preference for the control group
bc <- subset(BB, GROUP == "C")

bta <- subset(bc, PREF == "A")
btk <- subset(bc, PREF == "K")
btn <- subset(bc, PREF == "N")
btt <- subset(bc, PREF == "T")
btv <- subset(bc, PREF == "V")


dataAKNTV <- data.frame(group = c(rep("Auditory", length(bta$POST)), rep("Kinesthetic",length(btk$POST)), rep("No preference", length(btn$POST)), rep("Tactile", length(btt$POST)), rep("Visual", length(btv$POST))), value = c(bta$POST, btk$POST, btn$POST, btt$POST, btv$POST))

ggplot(dataAKNTV, aes(x = group, y = value, fill = group)) + geom_boxplot() + theme_classic()


# Data frame for each preference for the experiment group
be <- subset(BB, GROUP == "E")

bta <- subset(be, PREF == "A")
btk <- subset(be, PREF == "K")
btn <- subset(be, PREF == "N")
btt <- subset(be, PREF == "T")
btv <- subset(be, PREF == "V")



dataAKNTV <- data.frame(group = c(rep("Auditory", length(bta$POST)), rep("Kinesthetic",length(btk$POST)), rep("No preference", length(btn$POST)), rep("Tactile", length(btt$POST)), rep("Visual", length(btv$POST))), value = c(bta$POST, btk$POST, btn$POST, btt$POST, btv$POST))

ggplot(dataAKNTV, aes(x = group, y = value, fill = group)) + geom_boxplot() + theme_classic()



btm <- subset(BB, GENDER == "M")
btf <- subset(BB, GENDER == "F")

# Combine the data into a data frame
dataMF <- data.frame(group = c(rep("Male", length(btm$POST)), rep("Female", length(btf$POST))), value = c(btm$POST, btf$POST))

# Create the box plot graph
ggplot(dataMF, aes(x = group, y = value, fill = group)) + geom_boxplot() + theme_classic()
mean(btf$POST)-mean(btm$POST)
mean(btf$PRE)-mean(btm$PRE)
mean(btm$POST)
```


We have a dataset of 87 students, with 8 variables. (qualitative and quatitative)
– GROUP: C = ‘Control’ (i.e. traditional teaching methods), E =
‘Experiment’ (i.e. incorporated learning styles);
– SESSION: students were randomly assigned to 1 of 4 sessions (denoted by
S1 to S4) within each GROUP;
– PREF: students’ learning style preference (T = ‘Tactile’, K =
‘Kinesthetic’, A = ‘Auditory’, V = ‘Visual’, N = ‘No preference’);
– GENDER: M = male, F = female;
– SLEVEL: school level (E = ‘Elementary, H = ‘High School’);
– PRE: pre-test score (out of 100);
– POST: post-test score (out of 100);
– ATT: attitude scale score (out of 60).

We want to know if the students in the experiment group who were able to choose a PREF scored higher in a POST test than the students who were in a control group. PREF is the response variable. Build a model in R to answer the question.
The client wants you to investigate his research hypothesis that students in training sessions that utilise a processing activity that matches the students’ perceptual learning style preferences will demonstrate greater long-term retention of content than students in a traditional setting that has not utilised that processing activity.
Build a model in R considering t


```{r}

data <- read_excel("Bteaching.xls")
# Create a new variable for matching learning style preference
data$MATCH <- ifelse(data$PREF == "N", "ALL", data$PREF)

# Split the data into Control and Experiment groups
control_data <- subset(data, GROUP == "C")
experiment_data <- subset(data, GROUP == "E")

# Calculate the mean post-test scores for each group and learning style preference
control_means <- aggregate(control_data$POST, by = list(control_data$MATCH), FUN = mean)
experiment_means <- aggregate(experiment_data$POST, by = list(experiment_data$MATCH), FUN = mean)

```


# References {.unnumbered}

[gtsummary](https://www.danieldsjoberg.com/gtsummary/)

[Stratified gtsummary tables](https://www.danieldsjoberg.com/gtsummary/reference/tbl_strata.html)
