---
title: "CASE 4 - Teaching"
author: Michael Bigler, Tania Loureiro, William Elkiess, Dakota Cuellar
  and Ilyana El Mendili, Liam Phan
date: "2023-03-10"
output: 
  rmdformats::downcute:
    code_folding: hide
    toc_level: 6
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE,comment = FALSE, error = FALSE,options(scipen=999))
rm(list = ls())

```

# Packages

```{r packages}
library(knitr) # html tables
library(gtsummary)
library(dplyr)
source(file="RFn_Plot-lmSim.R")
library(leaps)
```

# Data Loading

```{r data loading}
teaching <- readxl::read_excel('Bteaching.xls')
teaching <- data.frame(teaching)
teaching$GROUP <- factor(teaching$GROUP, level = c('C', 'E'), labels = c('Control', 'Experiment'))
teaching$GENDER <- factor(teaching$GENDER, level = c('M', 'F'), labels =  c('Male', 'Female'))
# Don't need this teaching$TAUGHT <- factor(ifelse(teaching$PREF == 'N' | teaching$GROUP == 'Control', 'TR', teaching$PREF), level = c('TR', 'A', 'T', 'K', 'V'), labels = c('Traditional', 'Auditory', 'Tactile', 'Kinesthetic', 'Visual'))
teaching$PREF <- factor(teaching$PREF, level = c('A', 'N', 'T', 'K', 'V'), labels = c('Auditory', 'No Preference', 'Tactile', 'Kinesthetic', 'Visual')) # get understandable lables
teaching$SLEVEL <- factor(teaching$SLEVEL, level = c('E', 'H'), labels = c('Elementary', 'Highschool'))
```

# Data exploratory

```{r exploratory}
kable(summary(teaching))
```

```{r}
teaching$index = 1:nrow(teaching)
tsibble::tsibble(teaching, index = index)

table1 <- tbl_summary(
  teaching,
  by = GROUP, # split table by group
  missing = "no" # don't list missing data separately
) |>
  add_n() |>  # add column with total number of non-missing observations
  add_p() |>  # test for a difference between groups
  modify_header(label = "**Variable**") |> # update the column header
  bold_labels()
table1
```

```{r}
tbl_strata <-
  teaching |>
  tbl_strata(
    strata = GROUP,
    .tbl_fun =
      ~ .x |>
        tbl_summary(by = SESSION, missing = "no") |>
      add_n() |>  # add column with total number of non-missing observations
      add_p() |>  # test for a difference between groups
      modify_header(label = "**Variable**") |> # update the column header
      bold_labels() ,
    .header = "**{strata}**, N = {n}"
  )
tbl_strata
```

## Dummy encoding

```{r}
library(fastDummies)
teaching <- dummy_columns(teaching, select_columns = c('GROUP', 'SESSION', 'PREF', 'GENDER', 'SLEVEL'), remove_selected_columns = TRUE, remove_first_dummy = TRUE)
```


# DID

```{r}
teaching$ID <- 1:nrow(teaching)
teaching_DID <- teaching %>% slice(rep(1:n(), each = 2))
teaching_DID$POST_ind <- rep(c(0,1),nrow(teaching)) 
teaching_DID$OUTCOME <- ifelse(teaching_DID$POST_ind == 1, teaching_DID$POST, teaching_DID$PRE)
teaching_DID <- subset(teaching_DID, select = -c(POST, PRE))
```


## DID without covariates

```{r}
DID <- lm(OUTCOME ~ POST_ind + GROUP_Experiment + GROUP_Experiment*POST_ind, teaching_DID)
summary(DID)
plot(DID)
plot.lmSim(DID)
```

There is a treatment effect but no difference in the beginning between the groups. 

## DID with covariates 

```{r}
mod <- lm(OUTCOME ~ ATT + GROUP_Experiment + SESSION_S2 + SESSION_S3 + SESSION_S4 + `PREF_No Preference` + PREF_Tactile + PREF_Kinesthetic + PREF_Visual + GENDER_Female + SLEVEL_Highschool + ID + POST_ind + GROUP_Experiment * POST_ind, teaching_DID)
summary(mod)
plot(mod)
plot.lmSim(mod)
```

Model seems fine. We check the gam for linearity (of ATT as it is the only numerical value)

```{r}
# Check gam
library(gam)
mod.gam <- gam(OUTCOME ~ lo(ATT) + GROUP_Experiment + SESSION_S2 + SESSION_S3 + SESSION_S4 + `PREF_No Preference` + PREF_Tactile + PREF_Kinesthetic + PREF_Visual + GENDER_Female + SLEVEL_Highschool + ID + POST_ind + GROUP_Experiment * POST_ind, teaching_DID, family = gaussian(link = 'identity'))
plot(mod.gam, se = T)
```

Gam is fine. Linearity is ok. 

### Selection of variables

```{r}
library(regclass)
sort(VIF(mod), decreasing = TRUE)
```

Shows which variables have higher importance but we don't use this in the end

##### forward

```{r}
mod_min <- lm(OUTCOME ~ 1, teaching_DID)
s <- step(mod_min, direction='forward', scope=formula(lm(OUTCOME ~ ATT + GROUP_Experiment + SESSION_S2 + SESSION_S3 + SESSION_S4 + `PREF_No Preference` + PREF_Tactile + PREF_Kinesthetic + PREF_Visual + GENDER_Female + SLEVEL_Highschool + ID + POST_ind + GROUP_Experiment * POST_ind, teaching_DID)))
s$call
```

##### backward

```{r}
s <- step(mod, direction = 'backward')
s
s$call
```

Selects the best model backwards which makes some variables significant.

##### Combination backward and forward

```{r}
s <- step(mod)
s
s$call
```

##### Model

Based on forward and backward selection as well as the combination  we fit the following model

```{r}
mod2 <- lm(formula = lm(formula = OUTCOME ~ ATT + GROUP_Experiment + SESSION_S3 + PREF_Visual + POST_ind + GROUP_Experiment:POST_ind, data = teaching_DID),
    data = teaching_DID)
summary(mod2)
plot(mod2)
plot.lmSim(mod2)
```

We see there is no difference between the groups in the beginning. There is a difference between the times (post and pre). There is a treatment effect (interaction  term). ATT higher better score. Session 3 seems to be better taught. It could be that visual preference has a negative effect. 

# Stat stuff

```{r include=FALSE}
library(readxl)
BB <- read_excel("Bteaching.xls")
bt <- read_excel("Bteaching.xls")
data <- read_excel("Bteaching.xls")
```

$$
We\ build\ a\ GLM\ model:\\ \\POST\ = GROUP + PREF + PRE
$$

```{r}
# Convert GROUP and PREF to factors
#data$GROUP <- factor(data$GROUP)
#data$PREF <- factor(data$PREF)
data <- read_excel("Bteaching.xls")
library(ggplot2)
data$GENDER <- ifelse(data$GENDER == "M", 0, 1)
data$GROUP <- ifelse(data$GROUP == "C", 0, 1)
data$SLEVEL <- ifelse(data$SLEVEL == "E", 0, 1)
data$SESSION <- ifelse(data$SESSION == "S1", 1, 
                        ifelse(data$SESSION == "S2", 2, 
                               ifelse(data$SESSION == "S3", 3, 
                                      ifelse(data$SESSION == "S4", 4, NA))))
data$PREF <- ifelse(data$PREF == "A", 1, 
                        ifelse(data$PREF == "K", 2, 
                               ifelse(data$PREF == "N", 3, 
                                      ifelse(data$PREF == "T", 4,
                                             ifelse(data$PREF == "V", 5, NA)))))

colnames(data)[1] <- "Group"
colnames(data)[2] <- "Session"
colnames(data)[3] <- "Preference"
colnames(data)[4] <- "Gender"
colnames(data)[5] <- "School Level"
colnames(data)[6] <- "Pre Test Score"
colnames(data)[7] <- "Post Test Score"
colnames(data)[8] <- "Attention Test Score"
require(corrplot)

#2B3433, #2F4F4E, #0D8387, #CBE4DE
#("#FF0000", "#FFFFFF", "#0000FF")
cor_mat <- cor(data[, 7], data[, -7], method = "spearman")
col <- colorRampPalette(c("#2B3433", "#FFFFFF", "#0D8387"))(25)
corrplot(cor_mat, col = col, tl.col = "black",
         tl.srt = 40, tl.cex = 1.2,
         title = "Correlation Matrix", mar = c(0, 0, 1, 0))
#I have this corrplot i want to order the correlation circles from highest to lowest correlation
# Select the 7th column of the data and compute the correlation matrix


```

```{r}
data <- read_excel("Bteaching.xls")

data$SLEVEL <- factor(data$SLEVEL, level = c('E', 'H'), labels = c('Elementary', 'Highschool'))
data$PREF <- factor(data$PREF, level = c('A', 'N', 'T', 'K', 'V'), labels = c('A', 'N', 'T', 'K', 'V'))
data$GROUP <- factor(data$GROUP)


modelGLMAAA <- glm(formula = POST ~ GROUP + PRE, data = data, family =gaussian(link = "identity"))
summary(modelGLMAAA)

```

```{r}
data <- read_excel("Bteaching.xls")
data$GENDER <- ifelse(data$GENDER == "M", 0, 1)
data$GROUP <- ifelse(data$GROUP == "C", 0, 1)
data$SESSION <- ifelse(data$SESSION == "S1", 1, 
                        ifelse(data$SESSION == "S2", 2, 
                               ifelse(data$SESSION == "S3", 3, 
                                      ifelse(data$SESSION == "S4", 4, NA))))
bn <- subset(data, PRE < 56)
bf <- subset(data, PRE > 57)
bn$SLEVEL <- factor(bn$SLEVEL, level = c('E', 'H'), labels = c('Elementary', 'Highschool'))
bn$PREF <- factor(bn$PREF, level = c('A', 'N', 'T', 'K', 'V'), labels = c('A', 'N', 'T', 'K', 'V'))
bn$GROUP <- factor(bn$GROUP)


bf$SLEVEL <- factor(bf$SLEVEL, level = c('E', 'H'), labels = c('Elementary', 'Highschool'))
bf$PREF <- factor(bf$PREF, level = c('A', 'N', 'T', 'K', 'V'), labels = c('A', 'N', 'T', 'K', 'V'))
bf$GROUP <- factor(bf$GROUP)

modelGLMbn <- glm(formula = POST ~  GROUP + PRE , data = bn, family =gaussian(link = "identity"))
summary(modelGLMbn)
modelGLMbf <- glm(formula = POST ~  GROUP + PRE , data = bf, family =gaussian(link = "identity"))
summary(modelGLMbf)
```

```{r}
data <- read_excel("Bteaching.xls")
data$GENDER <- ifelse(data$GENDER == "M", 0, 1)
data$GROUP <- ifelse(data$GROUP == "C", 0, 1)
data$SESSION <- ifelse(data$SESSION == "S1", 1, 
                        ifelse(data$SESSION == "S2", 2, 
                               ifelse(data$SESSION == "S3", 3, 
                                      ifelse(data$SESSION == "S4", 4, NA))))
bE <- subset(data, SLEVEL == "E")
bH <- subset(data, SLEVEL == "H")
bE$SLEVEL <- factor(bE$SLEVEL, level = c('E', 'H'), labels = c('Elementary', 'Highschool'))
bE$PREF <- factor(bE$PREF, level = c('A', 'N', 'T', 'K', 'V'), labels = c('A', 'N', 'T', 'K', 'V'))
bE$GROUP <- factor(bE$GROUP)


bH$SLEVEL <- factor(bH$SLEVEL, level = c('E', 'H'), labels = c('Elementary', 'Highschool'))
bH$PREF <- factor(bH$PREF, level = c('A', 'N', 'T', 'K', 'V'), labels = c('A', 'N', 'T', 'K', 'V'))
bH$GROUP <- factor(bH$GROUP)

modelGLMbE <- glm(formula = POST ~  GROUP + PRE, data = bE, family =gaussian(link = "identity"))
summary(modelGLMbE)
modelGLMbH<- glm(formula = POST ~  GROUP + PRE, data = bH, family =gaussian(link = "identity"))
summary(modelGLMbH)
```

```{r}
data <- read_excel("Bteaching.xls")

data$GENDER <- factor(data$GENDER)
data$GROUP <- factor(data$GROUP)
data$SLEVEL <- factor(data$SLEVEL)
data$SESSION <- factor(data$SESSION)
data$PREF <- factor(data$PREF)
#data$PREF1A <- ifelse(data$PREF == "A", 1, 10)
#data$PREF2K <- ifelse(data$PREF == "K", 2, 10)
#data$PREF3N <- ifelse(data$PREF == "N", 3, 10)
#data$PREF5V <- ifelse(data$PREF == "V", 5, 10)
#data$PREF4T <- ifelse(data$PREF == "T", 4, 10)

modelGLM2 <- glm(POST ~  PRE + PREF + GENDER, data = data, family =gaussian(link = "identity"))
summary(modelGLM2)
plot(modelGLM2)
```

```{r}
library(ggplot2)
btc <- subset(BB, GROUP == "C")
bte <- subset(BB, GROUP == "E")

# Combine the data into a data frame
dataCE <- data.frame(group = c(rep("Control", length(btc$POST)), rep("Experiment", length(bte$POST))), value = c(btc$POST, bte$POST))

# Create the box plot graph
ggplot(dataCE, aes(x = group, y = value, fill = group)) + geom_boxplot() + theme_classic()

# Combine the data into a data frame
dataCE <- data.frame(group = c(rep("Control", length(btc$POST)), rep("Experiment", length(bte$POST))), value = c(btc$POST, bte$POST))


# Data frame for each preference for the control group
bc <- subset(BB, GROUP == "C")

bta <- subset(bc, PREF == "A")
btk <- subset(bc, PREF == "K")
btn <- subset(bc, PREF == "N")
btt <- subset(bc, PREF == "T")
btv <- subset(bc, PREF == "V")


dataAKNTV <- data.frame(group = c(rep("Auditory", length(bta$POST)), rep("Kinesthetic",length(btk$POST)), rep("No preference", length(btn$POST)), rep("Tactile", length(btt$POST)), rep("Visual", length(btv$POST))), value = c(bta$POST, btk$POST, btn$POST, btt$POST, btv$POST))

ggplot(dataAKNTV, aes(x = group, y = value, fill = group)) + geom_boxplot() + theme_classic()


# Data frame for each preference for the experiment group
be <- subset(BB, GROUP == "E")

bta <- subset(be, PREF == "A")
btk <- subset(be, PREF == "K")
btn <- subset(be, PREF == "N")
btt <- subset(be, PREF == "T")
btv <- subset(be, PREF == "V")



dataAKNTV <- data.frame(group = c(rep("Auditory", length(bta$POST)), rep("Kinesthetic",length(btk$POST)), rep("No preference", length(btn$POST)), rep("Tactile", length(btt$POST)), rep("Visual", length(btv$POST))), value = c(bta$POST, btk$POST, btn$POST, btt$POST, btv$POST))

ggplot(dataAKNTV, aes(x = group, y = value, fill = group)) + geom_boxplot() + theme_classic()



btm <- subset(BB, GENDER == "M")
btf <- subset(BB, GENDER == "F")

# Combine the data into a data frame
dataMF <- data.frame(group = c(rep("Male", length(btm$POST)), rep("Female", length(btf$POST))), value = c(btm$POST, btf$POST))

# Create the box plot graph
ggplot(dataMF, aes(x = group, y = value, fill = group)) + geom_boxplot() + theme_classic()
mean(btf$POST)-mean(btm$POST)
mean(btf$PRE)-mean(btm$PRE)
mean(btm$POST)
```

We have a dataset of 87 students, with 8 variables. (qualitative and quatitative) -- GROUP: C = 'Control' (i.e. traditional teaching methods), E = 'Experiment' (i.e. incorporated learning styles); -- SESSION: students were randomly assigned to 1 of 4 sessions (denoted by S1 to S4) within each GROUP; -- PREF: students' learning style preference (T = 'Tactile', K = 'Kinesthetic', A = 'Auditory', V = 'Visual', N = 'No preference'); -- GENDER: M = male, F = female; -- SLEVEL: school level (E = 'Elementary, H = 'High School'); -- PRE: pre-test score (out of 100); -- POST: post-test score (out of 100); -- ATT: attitude scale score (out of 60).

We want to know if the students in the experiment group who were able to choose a PREF scored higher in a POST test than the students who were in a control group. PREF is the response variable. Build a model in R to answer the question. The client wants you to investigate his research hypothesis that students in training sessions that utilise a processing activity that matches the students' perceptual learning style preferences will demonstrate greater long-term retention of content than students in a traditional setting that has not utilised that processing activity. Build a model in R considering t

```{r}

data <- read_excel("Bteaching.xls")
# Create a new variable for matching learning style preference
data$MATCH <- ifelse(data$PREF == "N", "ALL", data$PREF)

# Split the data into Control and Experiment groups
control_data <- subset(data, GROUP == "C")
experiment_data <- subset(data, GROUP == "E")

# Calculate the mean post-test scores for each group and learning style preference
control_means <- aggregate(control_data$POST, by = list(control_data$MATCH), FUN = mean)
experiment_means <- aggregate(experiment_data$POST, by = list(experiment_data$MATCH), FUN = mean)

```

# References {.unnumbered}

[gtsummary](https://www.danieldsjoberg.com/gtsummary/)

[Stratified gtsummary tables](https://www.danieldsjoberg.com/gtsummary/reference/tbl_strata.html)
